Fast solvers for the high-order FEM simplicial de Rham complex \thanks{Submitted to the editors July XX, 2024. \funding{ PDB and PEF were supported by EPSRC grant EP/W026260/1. } }

### 2024-06-14 paper/paper.tex
Remark on avoding  the FDM eigenproblem for the second type interior DOFs at the end of the complex

\begin{abstract}
We present new high-order finite elements discretizing the $L^2$ de Rham
complex on triangular and tetrahedral meshes. The finite elements discretize
the same spaces as usual, but with different basis functions. They allow for
fast linear solvers based on static condensation and space decomposition
methods. The new elements build upon the definition of degrees of freedom
given in L.~Demkowicz, P.~Monk, L.~Vardapetyan, and W.~Rachowicz.~
\textit{De Rham diagram for $hp$ finite element spaces}. Comput.~Math.~Appl.,
39(7-8):29--38, 2000., and consist of integral moments
on an equilateral reference simplex with respect to a numerically computed
polynomial basis that is orthogonal in both the $L^2$- and
$H(\mathrm{d})$-inner products ($\mathrm{d} \in \{\mathrm{grad},
\mathrm{curl}, \mathrm{div}\}$). On the reference equilateral simplex, the
resulting stiffness matrix has diagonal interior block, and does not couple
together the interior and interface degrees of freedom. Thus, on the
reference simplex, the Schur complement resulting from elimination of
interior degrees of freedom is simply the interface block itself. This
sparsity is not preserved on arbitrary cells mapped from the reference cell.
Nevertheless, the interior-interface coupling is weak because it is only
induced by the geometric transformation. We devise a preconditioning
strategy by neglecting the interior-interface coupling. We precondition the
interface Schur complement with the interface block, and simply apply
point-Jacobi to precondition the interior block. The combination of this
approach with a space decomposition method on small subdomains constructed
around vertices, edges, and faces allows us to efficiently solve the
canonical Riesz maps in $H(\mathrm{grad})$, $H(\mathrm{curl})$, and
$H(\mathrm{div})$, at very high order. We empirically demonstrate iteration
counts that are robust with respect to the polynomial degree. 
\end{abstract}

@@

We discretize $\Ltwo$-de Rham complex in triangular and tetrahedral meshes.

@@

Recall the definition of the $\Hcurl$-confroming finite element spaces
\begin{equation}
V^1_{p, p_E}(K) = \{v\in [\P_p(K)]^d : v\cdot\bt \in \P_{p_E}(E) \}
\end{equation}
for $p_E=p-1$ we obtain the \Nedelec space of the first kind $\Ned_p(K)$, while
$p_E=p$ gives the \Nedelec space of the second kind $\NedTwo_p(K) = [\P_p(K)]^d$. 
These spaces enforce continuity of the tangential trace across mesh interfaces.

We define a basis for the dual of $V^1_{p, p_E}(K)$ as 
tangential moments along $E\in \text{edges}(\Delta^d)$,
\begin{equation}
   \ell^E_j(v) = (q_j, v\cdot \bt)_E, \quad q_j \in \P_{p_E}(E), 
\end{equation}
and for each sub-entity $S \in \text{faces}(\Delta^d) \cup \text{interior}(\Delta^d)$,
\begin{align}
   \ell^{S,0}_j(v) &= (\grad_S\phi^S_j, v)_{S}, \\
   \ell^{S,1}_j(v) &= (\curl_S\Phi^S_j, \curl_S v)_{S},
\end{align}
where $\curl_S$ is the tangential curl on $S$,
$\{\phi^S_j\}$ are the $H_0(\grad, S)$ bases constructed in \eqref{eq:hgrad-fdm},
and $\{\curl_S\Phi^S_j\}$ is a basis for $\curl_S \mathbb{B}_{p_S}(\curl, S)$
such that
\begin{equation} \label{eq:hcurl-fdm}
   (\curl_S\Phi^S_j, \curl_S\Phi^S_i)_{S} = \delta_{ij}, \quad
   (\Phi^S_j, \Phi^S_i)_{S} = \lambda_j\delta_{ij}, \quad
   \Phi^S_j\times\bn = 0 \text{ on }\partial S.
\end{equation}

@@

Recall the definition of the $\Hdiv$-confroming finite element spaces
\begin{equation}
V^2_{p, p_F}(K) = \{v\in [\P_p(K)]^d : v\cdot\bn \in \P_{p_F}(F) \}
\end{equation}
for $p_F=p-1$ we obtain the Raviart--Thomas space $\RT_p(K)$, while
$p_F=p$ gives the Brezzi--Douglas--Marini space $\BDM_p(K) = [\P_p(K)]^d$. 
These spaces enforce continuity of the normal trace across mesh interfaces.

We define a basis for the dual of $V^2_{p, p_F}(K)$ as normal moments on $F\in
\text{faces}(\Delta^d)$,
\begin{equation}
   \ell^F_j(v) = (q_j, v\cdot \bn)_F, \quad q_j \in \P_{p_F}(F),
\end{equation}
and on the $\text{interior}(\Delta^d) = K$: 
\begin{align}
   \ell^{K,0}_j(v) &= (\curl\Phi^K_j, v)_{K}, \\
   \ell^{K,1}_j(v) &= (\div\Psi^K_j, \div v)_{K},
\end{align}
where $\{\Phi^K_j\}$ are the $H_0(\curl, S)$ bases constructed in
\eqref{eq:hcurl-fdm}, and $\{\div\Psi^K_j\}$ is a basis for $\div
\mathbb{B}_p(\div, K)$ such that
\begin{equation} \label{eq:hdiv-fdm}
   (\div\Psi^K_j, \div\Psi^K_i)_K = \delta_{ij}, \quad
   (\Psi^K_j, \Psi^K_i)_{K} = \lambda_j\delta_{ij}, \quad
   \Psi^K_j\cdot\bn = 0 \text{ on }\partial K.
\end{equation}

>>>>

\begin{abstract}
We present new high-order finite elements discretizing the $L^2$ de Rham
complex on triangular and tetrahedral meshes. The finite elements discretize
the same spaces as usual, but with different basis functions. They allow for
fast linear solvers based on static condensation and space decomposition
methods. The new elements build upon the definition of degrees of freedom
given in L.~Demkowicz, P.~Monk, L.~Vardapetyan, and W.~Rachowicz.~
\textit{De Rham diagram for $hp$ finite element spaces}. Comput.~Math.~Appl.,
39(7-8):29--38, 2000., and consist of integral moments
on an equilateral reference simplex with respect to a numerically computed
polynomial basis that is orthogonal in both the $L^2$- and
$H(\mathrm{d})$-inner products ($\mathrm{d} \in \{\mathrm{grad},
\mathrm{curl}, \mathrm{div}\}$). On the reference equilateral simplex, the
resulting stiffness matrix has diagonal interior block, and does not couple
together the interior and interface degrees of freedom. Thus, on the
reference simplex, the Schur complement resulting from elimination of
interior degrees of freedom is simply the interface block itself. This
sparsity is not preserved on arbitrary cells mapped from the reference cell.
Nevertheless, the interior-interface coupling is weak because it is only
induced by the geometric transformation. We devise a preconditioning
strategy by neglecting the interior-interface coupling. We precondition the
interface Schur complement with the interface block, and simply apply
point-Jacobi to precondition the interior block. The combination of this
approach with a space decomposition method on small subdomains constructed
around vertices and edges allows us to efficiently solve the
canonical Riesz maps in $H(\mathrm{grad})$, $H(\mathrm{curl})$, and
$H(\mathrm{div})$, at very high order. We empirically demonstrate iteration
counts that are robust with respect to the polynomial degree. 
\end{abstract}

@@

We discretize $\Ltwo$-de Rham complex on triangular and tetrahedral meshes.

@@

Recall the definition of the $\Hcurl$-conforming finite element spaces
\begin{equation}
V^1_{p, p_E}(K) = \{v\in [\P_p(K)]^d : v\cdot\bt \in \P_{p_E}(E) \}
\end{equation}
for $p_E=p-1$ we obtain the \Nedelec space of the first kind $\Ned_p(K)$, while
$p_E=p$ gives the \Nedelec space of the second kind $\NedTwo_p(K) = [\P_p(K)]^d$. 
These spaces enforce continuity of the tangential trace across mesh interfaces.

We define a basis for the dual of $V^1_{p, p_E}(K)$ as 
tangential moments along $E\in \text{edges}(\Delta^d)$,
\begin{equation}
   \ell^E_j(v) = (q_j, v\cdot \bt)_E, \quad q_j \in \P_{p_E}(E), 
\end{equation}
and for each sub-entity $S \in \text{faces}(\Delta^d) \cup \text{interior}(\Delta^d)$,
\begin{align}
   \ell^{S,0}_j(v) &= (\grad_S\phi^S_j, v)_{S}, \\
   \ell^{S,1}_j(v) &= (\curl_S\Phi^S_j, \curl_S v)_{S}, \label{eq:hcurl-interior-dofs}
\end{align}
where $\curl_S$ is the tangential curl on $S$,
$\{\phi^S_j\}$ are the $H_0(\grad, S)$ bases constructed in \eqref{eq:hgrad-fdm},
and $\{\curl_S\Phi^S_j\}$ is a basis for $\curl_S \mathbb{B}_{p_S}(\curl, S)$
such that
\begin{equation} \label{eq:hcurl-fdm}
   (\curl_S\Phi^S_j, \curl_S\Phi^S_i)_{S} = \delta_{ij}, \quad
   (\Phi^S_j, \Phi^S_i)_{S} = \lambda_j\delta_{ij}, \quad
   \Phi^S_j\times\bn = 0 \text{ on }\partial S.
\end{equation}
On a 2D face $F$ (or the interior of a 2D cell),
it is not necessary to solve for $\{\Phi^F_j\}$, as the degrees of freedom
\eqref{eq:hcurl-interior-dofs} only involve the scalar-valued 
$\curl_F\Phi^F_j$ which spans $\P_{p_F-1}(F)\setminus\P_0(F)$,
hence the standard $\Ltwo$-orthogonal Dubiner basis may be used in place of $\curl_F\Phi^F_j$.

@@

Recall the definition of the $\Hdiv$-conforming finite element spaces
\begin{equation}
V^2_{p, p_F}(K) = \{v\in [\P_p(K)]^d : v\cdot\bn \in \P_{p_F}(F) \}
\end{equation}
for $p_F=p-1$ we obtain the Raviart--Thomas space $\RT_p(K)$, while
$p_F=p$ gives the Brezzi--Douglas--Marini space $\BDM_p(K) = [\P_p(K)]^d$. 
These spaces enforce continuity of the normal trace across mesh interfaces.

We define a basis for the dual of $V^2_{p, p_F}(K)$ as normal moments on $F\in
\text{faces}(\Delta^d)$,
\begin{equation}
   \ell^F_j(v) = (q_j, v\cdot \bn)_F, \quad q_j \in \P_{p_F}(F),
\end{equation}
and on the $\text{interior}(\Delta^d) = K$: 
\begin{align}
   \ell^{K,0}_j(v) &= (\curl\Phi^K_j, v)_{K}, \\
   \ell^{K,1}_j(v) &= (\div\Psi^K_j, \div v)_{K}, \label{eq:hdiv-interior-dofs}
\end{align}
where $\{\Phi^K_j\}$ are the $H_0(\curl, S)$ bases constructed in
\eqref{eq:hcurl-fdm}, and $\{\div\Psi^K_j\}$ is a basis for $\div
\mathbb{B}_p(\div, K)$ such that
\begin{equation} \label{eq:hdiv-fdm}
   (\div\Psi^K_j, \div\Psi^K_i)_K = \delta_{ij}, \quad
   (\Psi^K_j, \Psi^K_i)_{K} = \lambda_j\delta_{ij}, \quad
   \Psi^K_j\cdot\bn = 0 \text{ on }\partial K.
\end{equation}
It is not necessary to solve for $\{\Psi^K_j\}$, as the degrees of freedom
\eqref{eq:hdiv-interior-dofs} only involve $\div\Psi^K_j$, 
which spans $\P_{p-1}(K)\setminus\P_0(K)$. Hence the standard
$\Ltwo$-orthogonal Dubiner basis may be used in place of $\div\Psi^K_j$.

\subsection{Generally-mapped elements}

The dual bases described above yield a sparse basis for $V^k_p(K)$ only
if $K$ is equilateral.
